Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm:

### 2023-07-01 manuscript.Rmd
First working version producing most figures

---
title: "Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm: 
A registered report"
author: "Matthew H.C. Mak, Alice O’Hagan, Aidan J Horner, M Gareth Gaskell"
date: "Department of Psychology, University of York, UK"
output:
   bookdown::pdf_document2:
    keep_tex: true
    toc: false
    fig_caption: yes
header-includes:
  # - \usepackage{setspace}\doublespacing
  - \usepackage{caption}
bibliography: references.bib
link-citations: true
linkcolor: black
csl: open-biology.csl
---

@@

packagelist <- c('bookdown','knitr','markdown','osfr','plyr','tidyr','dplyr','emmeans','ggplot2','Rmisc','buildmer','lme4','lmerTest','stringr','effects','effectsize','ggdist')
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

@@

studied$Delay <- str_replace_all(studied$Delay, "control", "Immediate/Control")
studied$Delay <- str_replace_all(studied$Delay, "delay", "12-hour delay")
studied$Delay <- factor(studied$Delay, levels = c("Immediate/Control","12-hour delay"))
```

```{r intrusionsbasic}
intrusions <- distinct(lure, Participant.Private.ID, .keep_all = T)
intrusions$answer <- NULL
intrusions$original_response <- NULL
intrusions$lure_recalled <- NULL
```

```{r poissonregression}
intrusion_poisson <- glm(intrusions ~ Delay*Test.Time, data = intrusions, family = poisson(link = "log"))

intrusionout <- emmeans(intrusion_poisson, pairwise ~ Test.Time | Delay)
#Test time (PM>AM) showed a large effect on output bias
```

@@

pdf('Figures/intrusiongraph.pdf',width=15,height=12)

ggplot(intrusions%>%
         mutate(Test.Time = case_when(Delay=="Immediate/Control"~Test.Time,
                       Test.Time== "AM" ~ "Sleep",
                       TRUE ~ "Wake")) ,
       aes(Test.Time, intrusions, fill=Test.Time))+
  ggdist::stat_halfeye(adjust = .9, width = .7, .width = 0, justification = -.2, point_colour = NA, alpha=0.5) + 
  stat_summary(fun=mean, geom = "bar", width=0.3)+
  labs(y="Number of Intrusions", x="", title="Intrusion", size=30)+
  geom_errorbar(width=.1, aes(ymin=intrusions-ci, ymax=intrusions+ci), position = position_dodge(width = 0.2), data=intrusion_CI%>%
         mutate(Test.Time = case_when(Delay=="Immediate/Control"~Test.Time,
                       Test.Time== "AM" ~ "Sleep",
                       TRUE ~ "Wake")))+
  scale_y_continuous(expand = c(0,0),limits = c(0,60),breaks=seq(0,60,5))+
  theme_light()+
  theme(legend.position="none",legend.title = element_blank(), axis.title.y = element_text(size=16),plot.title = element_text(size=25, face="bold"), axis.text.x = element_text(size=16),axis.text.y = element_text(size=25),legend.text = element_text(size=25), axis.title=element_text(size=25,face="bold"),strip.text = element_text(size = 25), aspect.ratio =1/1)+ scale_fill_manual(values = c("darkorange", "cornflowerblue", "darkorange", "cornflowerblue" ))+coord_cartesian(ylim = c(0,60))+facet_wrap(~Delay,scales = "free_x")+ geom_dotplot(binaxis = "y", stackdir = "down",dotsize = 0.25)

@@

@@

To address Research Question #1, we first tested if any difference between the Sleep and Wake groups was significantly different from that between the AM- and PM-control groups (i.e., an interaction between Interval and Test Time). This is important because it allows us to rule out time-of-day effects. Then, we tested for the simple effect of Test Time (9AM vs. 9PM) within the Immediate and Delay groups. If there is (1) a significant Interval $\times$ Test Time interaction and (2) a significant Test Time effect within the Delay groups (Sleep > Wake), we will be able conclude that sleep (but not time-of-day) increases false recall.

@@

The procedure of the study is summarised in Figure 1. The study was hosted on Gorilla (www.gorilla.sc; [@anwyl-irvine2020a]). A study phase took approximately 11 minutes. Here, participants first gave informed consent, completed a language/attention check, rated their level of sleepiness on the Stanford Sleepiness Scale (SSS; [@hoddes1973a]), and viewed 20 DRM wordlists.

@@

@@

During presentation, words in each DRM list were presented visually,   in a fixed order and arranged in descending associative strength to the unpresented critical lure (see Table 2 for order). Each list began with a fixation for 1 s, followed by the first word in a list. Each word was shown for 1 s, in a lowercase black font (Arial, size 26) on a white background, and separated by a 500-ms interstimulus interval. After presentation of the final word in a list was 5 s of blank screen. List order was randomised, and each list was seen once.

@@

The free recall data were pre-processed. The first step was to remove any duplicate responses. The second was to correct all obvious spelling and typing errors to the nearest English words, defined as Levenstein distance $\leq$ 2 (e.g., *cigerette → cigarette).   Responses with added or dropped inflectional suffixes (i.e., -s, -ed, -ing, adjectival -er) were corrected. Responses with derivational changes were considered as intrusions. For instance, one of the studied words is pollution; if a participant recalled pollutions, the plural suffix was dropped; however, if a participant recalled pollutant, this was considered as an intrusion. 
# Results of Pre-registered Analyses

@@

## Control analysis.
Payne et al. [-@payne2009a] found that participants falsely recalled more critical lures post-sleep (vs. post-wake). However, it is possible that participants simply had a greater tendency to put down more unseen words after sleep, not because sleep increases DRM false recall per se. Therefore, before addressing our key research questions, we checked if participants across groups were comparable in terms of their bias in producing unseen items. In Payne et al. [-@payne2009a], this bias was indexed via the number of intrusions (i.e., neither the studied nor the lure items), which was roughly equivalent between their Sleep and Wake groups [MSleep = 5.6 vs. MWake = 6.2; p = .60] as well as between their AM and PM-control groups [MAM = 4.1 vs. MPM = 4.1; p = .99]. To check if this is the case in our data, we used a 2 (Interval: Immediate vs. Delay) $\times$ 2 (Test Time: AM vs. PM) Poisson regression. We chose Poisson regression, as opposed to ANOVA, because the intrusion data were count data, meaning that data distribution were right-skewed and hence unsuitable for ANOVA.  Figure 2 summarises the number of intrusions in each group.

@@

The number of critical lures falsely recalled is summarised across groups in Figure 3.

>>>>

---
title: "Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm: 
A registered report"
author: "Matthew H.C. Mak, Alice O’Hagan, Aidan J. Horner, M. Gareth Gaskell"
date: "Department of Psychology, University of York, UK"
output:
   bookdown::pdf_document2:
    keep_tex: true
    toc: false
    fig_caption: yes
header-includes:
  # - \usepackage{setspace}\doublespacing
  - \usepackage{caption}
bibliography: references.bib
link-citations: true
linkcolor: black
csl: proceedings-of-the-royal-society-b.csl
---

@@

processdata <- 0   # running all analyses from scratch takes about 6 hours
buildfigures <- 0

packagelist <- c('bookdown','knitr','markdown','osfr','plyr','tidyr','dplyr','emmeans','ggplot2','Rmisc','buildmer','lme4','lmerTest','stringr','effects','effectsize','ggdist','car','gridExtra')
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

if (processdata==0){load('data/processeddata.RData')}

@@

studied$Delay <- str_replace_all(studied$Delay, "control", "Immediate/Control")
studied$Delay <- str_replace_all(studied$Delay, "delay", "12-hour delay")
studied$Delay <- factor(studied$Delay, levels = c("Immediate/Control","12-hour delay"))

intrusions <- distinct(lure, Participant.Private.ID, .keep_all = T)
intrusions$answer <- NULL
intrusions$original_response <- NULL
intrusions$lure_recalled <- NULL
```

```{r poissonregression}

if (processdata==1){
intrusion_poisson <- glm(intrusions ~ Delay*Test.Time, data = intrusions, family = poisson(link = "log"))

intrusionout <- emmeans(intrusion_poisson, pairwise ~ Test.Time | Delay)
#Test time (PM>AM) showed a large effect on output bias
}

@@

if (buildfigures==1){
pdf('Figures/intrusiongraph.pdf',width=15,height=10)

plothandle <- ggplot(intrusions%>%
         mutate(Test.Time = case_when(Delay=="Immediate/Control"~Test.Time,
                       Test.Time== "AM" ~ "Sleep",
                       TRUE ~ "Wake")) ,
       aes(Test.Time, intrusions, fill=Test.Time))+
  ggdist::stat_halfeye(adjust = .9, width = .7, .width = 0, justification = -.2, point_colour = NA, alpha=0.5) + 
  stat_summary(fun=mean, geom = "bar", width=0.3)+
  labs(y="Number of Intrusions", x="", title="Intrusion", size=30)+
  geom_errorbar(width=.1, aes(ymin=intrusions-ci, ymax=intrusions+ci), position = position_dodge(width = 0.2), data=intrusion_CI%>%
         mutate(Test.Time = case_when(Delay=="Immediate/Control"~Test.Time,
                       Test.Time== "AM" ~ "Sleep",
                       TRUE ~ "Wake")))+
  scale_y_continuous(expand = c(0,0),limits = c(0,60),breaks=seq(0,60,5))+
  theme_light()+
  theme(legend.position="none",legend.title = element_blank(), axis.title.y = element_text(size=16),plot.title = element_text(size=25, face="bold"), axis.text.x = element_text(size=16),axis.text.y = element_text(size=25),legend.text = element_text(size=25), axis.title=element_text(size=25,face="bold"),strip.text = element_text(size = 25), aspect.ratio =1/1)+ scale_fill_manual(values = c("darkorange", "cornflowerblue", "darkorange", "cornflowerblue" ))+coord_cartesian(ylim = c(0,60))+facet_wrap(~Delay,scales = "free_x")+ geom_dotplot(binaxis = "y", stackdir = "down",dotsize = 0.25)

grid.arrange(plothandle, ncol=1)

dev.off()
}

```{r lurecontrast}
lure$Delay <- as.factor(lure$Delay)
contrasts(lure$Delay) <- contr.sum(2)   
lure$Test.Time <- as.factor(lure$Test.Time)
contrasts(lure$Test.Time) <- contr.sum(2)
```

```{r lureGLM}

if (processdata==1){
lure_model_buildmer <- buildmer(lure_recalled ~ Delay * Test.Time +intrusions +
                                       (intrusions | Participant.Private.ID) + 
                                       (Delay * Test.Time| answer), 
                                     data = lure, 
                                     family = "binomial",
                                     buildmerControl = buildmerControl(direction='backward',args =list(control=glmerControl(optimizer="bobyqa"))))

lure_model <- glmer(formula(lure_model_buildmer), data = lure, family = binomial(link = 'logit'), control = glmerControl(optimizer = 'bobyqa'))

# pre-registered pair-wise comparison
lure_pairwise <- emmeans(lure_model, pairwise ~ Test.Time | Delay)

# exploratory analysis to obtain an effect size for the effect of sleep
lureeffects <- eff_size(lure_pairwise, sigma(lure_model), edf = Inf) 
}

```{r lureGLMnointrusions}
if (processdata==1){
lure_model_without_covariate_buildmer <- buildmer(lure_recalled ~ Delay * Test.Time +
                                       (1 | Participant.Private.ID) + 
                                       (Delay * Test.Time| answer), 
                                     data = lure, 
                                     family = "binomial",
                                     buildmerControl = buildmerControl(direction='backward',args =list(control=glmerControl(optimizer="bobyqa"))))
}

```{r lureBayesian}

if (processdata==1){
null_lure <- update(lure_model, formula = ~ . -Delay*Test.Time)

#Bayes Factor
lureBF <- exp((BIC(null_lure)-BIC(lure_model))/2)
}

```{r lureBayesianSE}

if (processdata==1){
#simple effects of test time
lure_immediate <- subset(lure, Delay=="Immediate/Control")

lure_delay <- subset(lure, Delay=="12-hour delay")

#delay
lure_delay_model1 <- buildmer(lure_recalled ~ Test.Time +intrusions +
                                       (intrusions | Participant.Private.ID) + 
                                       (Test.Time| answer), 
                                     data = lure_delay, 
                                     family = "binomial",
                                     buildmerControl = buildmerControl(direction='backward',args =list(control=glmerControl(optimizer="bobyqa"))))

lure_delay_model1_null <- update(lure_delay_model1, formula = ~ . -Test.Time)

luredelayBF <- exp(BIC(lure_delay_model1_null)-BIC(lure_delay_model1))/2 #= 0.001937424

#immediate
lure_immediate_model1 <- buildmer(lure_recalled ~ Test.Time +intrusions +
                                       (intrusions | Participant.Private.ID) + 
                                       (Test.Time| answer), 
                                     data = lure_immediate, 
                                     family = "binomial",
                                     buildmerControl = buildmerControl(direction='backward',args =list(control=glmerControl(optimizer="bobyqa"))))

lure_immediate_model1_null <- update(lure_immediate_model1, formula = ~ . -Test.Time)

lureimmediateBF <- exp(BIC(lure_immediate_model1_null)-BIC(lure_immediate_model1))/2 #= 0.0001069749
}

```{r lureposcontrol}

lure_pp <- ddply(lure, .(Delay, Test.Time, Participant.Private.ID), summarise, Lure.Recalled=sum(lure_recalled))

if (processdata==1){

#one sample t-test
lurettest <- t.test(lure_pp$Lure.Recalled, mu=0) #t = 26.964, df = 487, p-value < 2.2e-16

}

```{r makeluregraph}
#confidence interval for graph
lure_CI <- Rmisc::summarySE(lure_pp, measurevar = "Lure.Recalled", groupvars = c("Delay", "Test.Time"), conf.interval = 0.95)

if (buildfigures==1){
pdf('Figures/luregraph.pdf',width=15,height=10)

plothandle <- ggplot(lure_pp%>%
         mutate(Test.Time = case_when(Delay=="Immediate/Control"~Test.Time,
                       Test.Time== "AM" ~ "Sleep",
                       TRUE ~ "Wake")) ,
       aes(Test.Time, Lure.Recalled, fill=Test.Time))+
  ggdist::stat_halfeye(adjust = .9, width = .7, .width = 0, justification = -.2, point_colour = NA, alpha=0.5) + 
  stat_summary(fun=mean, geom = "bar", width=0.3)+
  labs(y="Number of Lures Falsely Recalled (Max = 20)", x="", title="Lure", size=30)+
  geom_errorbar(width=.1, aes(ymin=Lure.Recalled-ci, ymax=Lure.Recalled+ci), position = position_dodge(width = 0.2), data=lure_CI%>%
         mutate(Test.Time = case_when(Delay=="Immediate/Control"~Test.Time,
                       Test.Time== "AM" ~ "Sleep",
                       TRUE ~ "Wake")))+
  scale_y_continuous(expand = c(0,0),limits = c(-0.1,15.5),breaks=seq(0,15,2))+
  theme_light()+
  theme(legend.position="none",legend.title = element_blank(), axis.title.y = element_text(size=16),plot.title = element_text(size=25, face="bold"), axis.text.x = element_text(size=16),axis.text.y = element_text(size=25),legend.text = element_text(size=25), axis.title=element_text(size=25,face="bold"),strip.text = element_text(size = 25), aspect.ratio =1/1)+ scale_fill_manual(values = c("darkorange", "cornflowerblue", "darkorange", "cornflowerblue" ))+coord_cartesian(ylim = c(-0.1,15.1))+facet_wrap(~Delay,scales = "free_x")+ geom_dotplot(binaxis = "y", stackdir = "down",dotsize = 0.25)

grid.arrange(plothandle, ncol=1)

dev.off()
}

```{r studiededits}

studied$Delay <- as.factor(studied$Delay)
contrasts(studied$Delay) <- contr.sum(2)   
studied$Test.Time <- as.factor(studied$Test.Time)
contrasts(studied$Test.Time) <- contr.sum(2)

```{r studiedGLM}

if (processdata==1){
studied_model_buildmer <- buildmer(studied_recalled ~ Delay * Test.Time +intrusions + (intrusions | Participant.Private.ID) + (Delay * Test.Time| answer), data = studied, family = "binomial",buildmerControl = buildmerControl(direction='backward', args = list(control=glmerControl(optimizer="bobyqa"))))

studied_model <- glmer(formula(studied_model_buildmer), data = studied, family = binomial(link = 'logit'), control = glmerControl(optimizer = 'bobyqa'))

# studied_model <- glmer(formula(studied_model_buildmer), data = studied, family = binomial(link = 'logit'), nAGQ=0, control = glmerControl(optimizer = 'nloptwrap'))

studiedem1 <- emmeans(studied_model, pairwise ~ Test.Time | Delay)
studiedem2 <- emmeans(studied_model, pairwise ~ Delay | Test.Time)
}

```{r studiedBF}
if (processdata==1){
null_studied <- update(studied_model, formula = ~ . -Delay*Test.Time)

#Bayes Factor
studiedBF <- exp((BIC(null_studied)-BIC(studied_model))/2) #29122036
}

```{r studiedBayesian}

if (processdata==1){
#simple effects of test time
studied_immediate <- subset(studied, Delay=="Immediate/Control")

studied_delay <- subset(studied, Delay=="12-hour delay")

#delay
studied_delay_model1 <- buildmer(studied_recalled ~ Test.Time +intrusions +
                                       (intrusions | Participant.Private.ID) + 
                                       (Test.Time| answer), 
                                     data = studied_delay, 
                                     family = "binomial",
                                     buildmerControl = buildmerControl(direction='backward',args =list(control=glmerControl(optimizer="bobyqa"))))

studied_delay_model1_null <- update(studied_delay_model1, formula = ~ . -Test.Time)

studdelayBF <- exp(BIC(studied_delay_model1_null)-BIC(studied_delay_model1))/2 #0.01763277

#immediate
studied_immediate_model1 <- buildmer(studied_recalled ~ Test.Time +intrusions +
                                       (intrusions | Participant.Private.ID) + 
                                       (Test.Time| answer), 
                                     data = studied_immediate, 
                                     family = "binomial",
                                     buildmerControl = buildmerControl(direction='backward',args =list(control=glmerControl(optimizer="bobyqa"))))

studied_immediate_model1_null <- update(studied_immediate_model1, formula = ~ . -Test.Time)

studimBF <- exp(BIC(studied_immediate_model1_null)-BIC(studied_immediate_model1))/2 #0.0003863469

}
```

```{r makestudiedgraph}
studied_pp <- ddply(studied, .(Delay, Test.Time, Participant.Private.ID), summarise, Studied.Recalled=sum(studied_recalled))
studied_CI <- Rmisc::summarySE(data = studied_pp, measurevar = "Studied.Recalled", groupvars = c("Delay", "Test.Time"), conf.interval = 0.95)

if (buildfigures==1){
  pdf('Figures/studiedgraph.pdf',width=15,height=10)

plothandle <- ggplot(studied_pp%>%
         mutate(Test.Time = case_when(Delay=="Immediate/Control"~Test.Time,
                       Test.Time== "AM" ~ "Sleep",
                       TRUE ~ "Wake")) ,
       aes(Test.Time, Studied.Recalled, fill=Test.Time))+
  ggdist::stat_halfeye(adjust = .9, width = .7, .width = 0, justification = -.2, point_colour = NA, alpha=0.5) + 
  stat_summary(fun=mean, geom = "bar", width=0.3)+
  labs(y="Number of Studied Words Recalled (Max = 160)", x="", title="Studied", size=30)+
  geom_errorbar(width=.1, aes(ymin=Studied.Recalled-ci, ymax=Studied.Recalled+ci), position = position_dodge(width = 0.2), data=studied_CI%>%
         mutate(Test.Time = case_when(Delay=="Immediate/Control"~Test.Time,
                       Test.Time== "AM" ~ "Sleep",
                       TRUE ~ "Wake")))+
  scale_y_continuous(expand = c(0,0),limits = c(0,110),breaks=seq(0,110,10))+
  theme_light()+
  theme(legend.position="none",legend.title = element_blank(), axis.title.y = element_text(size=16),plot.title = element_text(size=25, face="bold"), axis.text.x = element_text(size=16),axis.text.y = element_text(size=25),legend.text = element_text(size=25), axis.title=element_text(size=25,face="bold"),strip.text = element_text(size = 25), aspect.ratio =1/1)+ scale_fill_manual(values = c("darkorange", "cornflowerblue", "darkorange", "cornflowerblue" ))+coord_cartesian(ylim = c(0,110))+facet_wrap(~Delay,scales = "free_x")+ geom_dotplot(binaxis = "y", stackdir = "down",dotsize = 0.25)

grid.arrange(plothandle, ncol=1)

@@

}
```

```{r excludeparticipants}

raw_fm <- raw_fm %>% separate(Experiment.ID, c("Delay", "Test.Time"))

fm <- subset(raw_fm,  subset = !Participant.Private.ID %in% c(7862197,
7636967,
7733905,
7643108,
7848363,
8307220,
8280100,
8401290,
8354767,
8209020,
7674684,
7814523,
7954183,
8138170,
8246649,
8393638,
7968031,
7806505,
8137770,
8116661,
8190616,
7655503,
7818063,
7999266,
8200792,
8225383,
8394785,
7817900,
7839745,
8201245,
8192060,
7941696,
7657171,
8201078,
8141329,
7676682,
8007863,
8021105,
7649898,
8171977,
8345258,
8061284,
8196008,
8471724,
8511638,
8474922,
8474259))

fm2 <- fm
fm2$response[fm2$response == ""] <- NA

fm2$filled <- ifelse(is.na(fm2$response), 0, 1)

total <- ddply(fm2, .(Participant.Private.ID, Delay, Test.Time), summarise, total=sum(filled))

# total response per group (Mean and SD)
ddply(total, .(Delay, Test.Time), summarise, average=mean(total), stdev=sd(total))

# ANOVA
total_anova <- aov(log10(total) ~ Delay*Test.Time, data = total)
total_resid <- car::qqPlot(total_anova$residuals, id = FALSE) #looks great

```{r recalleffect}

#just some merging for easier manipulation
studied_pp2 <- select(studied_pp, Participant.Private.ID, Studied.Recalled)
intrusions2 <- select(intrusions, Participant.Private.ID, intrusions)
lure_pp_explore <- left_join(lure_pp, studied_pp2, by="Participant.Private.ID")
lure_pp_explore <- left_join(lure_pp_explore, intrusions2,by="Participant.Private.ID")
lure_pp_explore_delay <- subset(lure_pp_explore, Delay=="12-hour delay")

#obtain adjusted recall (correct recall - intrusion)
lure_pp_explore_delay$adjusted_recall <- lure_pp_explore_delay$Studied.Recalled-lure_pp_explore_delay$intrusions
median(lure_pp_explore_delay$adjusted_recall) # adjusted recall median = 5
lure_pp_explore_delay$adjusted_recall_binary <- lure_pp_explore_delay$adjusted_recall
lure_pp_explore_delay$adjusted_recall_binary <- ifelse(lure_pp_explore_delay$adjusted_recall_binary>5, "High performer", "Low performer")

ddply(lure_pp_explore_delay, .(Test.Time,adjusted_recall_binary), summarise, average=mean(Lure.Recalled), stdev=sd(Lure.Recalled))

#poisson regression
prsum <- summary(lure_explore <- glm(Lure.Recalled ~ Test.Time*adjusted_recall_binary, data = lure_pp_explore_delay, family = poisson(link = "log")))

prem <- emmeans(lure_explore, pairwise ~ Test.Time | adjusted_recall_binary)

#graph (not reported in manuscript)
#lure_pp_explore_delay$Test.Time <- str_replace_all(lure_pp_explore_delay$Test.Time, "AM", "Sleep")
#lure_pp_explore_delay$Test.Time <- str_replace_all(lure_pp_explore_delay$Test.Time, "PM", "Wake")
#ggplot(lure_pp_explore_delay, aes(Test.Time,Lure.Recalled, fill=Test.Time))+stat_summary(fun=mean, geom = "bar")+  stat_summary(fun=mean, geom = "bar", width=0.01,position =position_dodge(width=.01))+scale_y_continuous(expand = c(0,0),limits = c(0,20),breaks=seq(0,20,2))+theme_light()+theme(legend.position="none",legend.title = element_blank(), axis.title.y = element_text(size=16),plot.title = element_text(size=25, face="bold"), axis.text.x = element_text(size=16),axis.text.y = element_text(size=25),legend.text = element_text(size=25), axis.title=element_text(size=25,face="bold"),strip.text = element_text(size = 25), aspect.ratio =1/1)+ scale_fill_manual(values = c("darkorange", "cornflowerblue"))+coord_cartesian(ylim = c(0,20))+ geom_dotplot(binaxis = "y", stackdir = "down",dotsize = 0.25)+facet_wrap(~adjusted_recall_binary)

```{r dependency}

studied2 <- studied

studied2 <- left_join(studied2, material, by="answer")

studied2_by_lure <- ddply(studied2, .(Participant.Private.ID, lure), summarise, N_per_list = sum(studied_recalled))

lure2 <- lure

lure2 <- rename(lure2, lure = answer)

lure2 <- select(lure2, Participant.Private.ID, Delay, Test.Time, lure, lure_recalled, intrusions)

iota_explore <- left_join(studied2_by_lure, lure2, by=c("Participant.Private.ID", "lure"))

#delay only
iota_explore_delay <- subset(iota_explore, Delay=="12-hour delay")

#wake group only
iota_explore_delay_wake <- subset(iota_explore_delay, Test.Time=="PM")

#sleep group only
iota_explore_delay_sleep <- subset(iota_explore_delay, Test.Time=="AM")

```{r allgroups}

if (processdata==1){
iota_buildmer000 <- buildmer(lure_recalled ~ Delay*Test.Time*N_per_list+ intrusions +
                               (N_per_list+ intrusions | Participant.Private.ID) + 
                               (Test.Time| lure), 
                           data = iota_explore, 
                           family = "binomial",
                           buildmerControl = buildmerControl(direction='backward',args =list(control=glmerControl(optimizer="bobyqa"))))

}

#Given the significant Delay1:Test.Time1:N_per_list, we looked at the Immediate and Delay groups separately below
```

```{r delaygroup}
if (processdata){
iota_buildmer2 <- buildmer(lure_recalled ~ Test.Time*N_per_list+ intrusions +
                                       (N_per_list+ intrusions | Participant.Private.ID) + 
                                       (Test.Time| lure), 
                                     data = iota_explore_delay, 
                                     family = "binomial",
                                     buildmerControl = buildmerControl(direction='backward',args =list(control=glmerControl(optimizer="bobyqa"))))

iota_buildmer2_standard <- glmer(formula(iota_buildmer2), data = iota_explore_delay, family = binomial(link = 'logit'), control = glmerControl(optimizer = 'bobyqa'))

}

```{r immediategroup}

if (processdata){
iota_explore_immediate <- subset(iota_explore, Delay=="Immediate/Control")

iota_buildmer3_standard <- glmer(lure_recalled ~Test.Time*N_per_list+intrusions+
                                       (1 | Participant.Private.ID) + 
                                       (1| lure), data = iota_explore_immediate, family = binomial(link = 'logit'), control = glmerControl(optimizer = 'bobyqa'))

save(file='data/processeddata.RData',list=c('intrusion_poisson','intrusionout','intrusion_CI','lure_model_buildmer','lure_model','lure_pairwise','lureeffects','lure_model_without_covariate_buildmer','null_lure','lureBF','lure_delay_model1','lure_immediate_model1','lure_delay_model1_null','lure_immediate_model1_null','luredelayBF','lureimmediateBF','lurettest','studied_model_buildmer','studied_model','studiedem1','studiedem2','studiedBF','studied_delay_model1','studied_delay_model1_null','studdelayBF','studied_immediate_model1','studied_immediate_model1_null','studimBF','prem','prsum','iota_buildmer000','iota_buildmer2','iota_buildmer2_standard','iota_buildmer3_standard'))

}

```{r makedelaygraph}

eff.p1 <- effect("Test.Time:N_per_list", iota_buildmer2_standard, KR=T)
eff.p1 <- as.data.frame(eff.p1)
eff.p1$Test.Time <- ifelse(eff.p1$Test.Time=="AM", "Sleep", "Wake")

probabilities <- ddply(iota_explore_delay, .(Participant.Private.ID, Test.Time), summarise, N_per_list_mean=mean(N_per_list), lure_mean=mean(lure_recalled))

if (buildfigures==1){

pdf('Figures/delaygraph.pdf',width=18,height=9)

plot1 <- ggplot(eff.p1, aes(N_per_list, color = Test.Time)) +
  geom_line(aes(y = fit, group=factor(Test.Time)), size=1.2) +
  geom_line(aes(y = lower,
                group=factor(Test.Time)), linetype =3) +
  geom_line(aes(y = upper,
                group=factor(Test.Time)), linetype =3) +
  xlab("Correct recall per list (Max = 8)") +
  ylab("Lure recall probability") +
  scale_colour_manual(values = c("darkorange", "cornflowerblue"))+
  labs(title="Prediction from mixed-effect model",color='Test.Time') + theme_bw()+   theme(legend.title = element_blank(), axis.title.y = element_text(size=20),plot.title = element_text(size=20, face="bold"), axis.text.x = element_text(size=20),axis.text.y = element_text(size=20),legend.text = element_text(size=20), axis.title=element_text(size=20,face="bold"),strip.text = element_text(size = 20), aspect.ratio =1/1)+coord_cartesian(xlim = c(0,8), ylim = c(0,1))+ scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))

plot2 <- ggplot(probabilities, aes(N_per_list_mean, lure_mean, colour=Test.Time))+geom_point() + geom_smooth(method = "lm", se = TRUE, aes(fill=Test.Time))+ xlab("Mean correct recall per list (Max = 8)") +
    ylab("Mean lure recall probability") +
    scale_colour_manual(values = c("darkorange", "cornflowerblue"))+
    labs(title='Actual data') + theme_bw()+   theme(legend.title = element_blank(), axis.title.y = element_text(size=20),plot.title = element_text(size=20, face="bold"), axis.text.x = element_text(size=20),axis.text.y = element_text(size=20),legend.text = element_text(size=20), axis.title=element_text(size=20,face="bold"),strip.text = element_text(size = 20), aspect.ratio =1/1)+coord_cartesian(xlim = c(0,8), ylim = c(0,1))+ scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))

grid.arrange(plot1, plot2, ncol=2)

}

@@

To address Research Question \#1, we first tested if any difference between the Sleep and Wake groups was significantly different from that between the AM- and PM-control groups (i.e., an interaction between Interval and Test Time). This is important because it allows us to rule out time-of-day effects. Then, we tested for the simple effect of Test Time (9AM vs. 9PM) within the Immediate and Delay groups. If there is (1) a significant Interval x Test Time interaction and (2) a significant Test Time effect within the Delay groups (Sleep > Wake), we will be able conclude that sleep (but not time-of-day) increases false recall.\footnote{Prior studies in the ‘Sleep $\times$ DRM’ literature (e.g., Fenn et al., 2009; Payne et al., 2009) conducted two separate statistical tests, one comparing Sleep vs. Wake, another comparing AM- vs. PM- controls. They then concluded that Sleep had an effect on DRM false memory beyond time-of-day effects when Test Time was significant in the Sleep vs. Wake comparison ($p$ < 0.05) but not in the AM vs. PM comparison ($p$ > 0.05). Unfortunately, however, this is not sufficient (Nieuwenhuis et al., 2011), as “the difference between ‘significant’ and ‘not significant’ is not itself statistically significant” (Gelmam \& Stern, 2006). Therefore, in order to rule out time-of-day effects, one needs to show that Sleep vs. Wake is significantly different from AM vs. PM-control. This can be captured by an Interval $\times$ Test Time interaction.}

@@

The procedure of the study is summarised in Figure \@ref(fig:expprocedure). The study was hosted on Gorilla (www.gorilla.sc; [@anwyl-irvine2020a]). A study phase took approximately 11 minutes. Here, participants first gave informed consent, completed a language/attention check, rated their level of sleepiness on the Stanford Sleepiness Scale (SSS; [@hoddes1973a]), and viewed 20 DRM wordlists.

@@

```{r expprocedure, fig.cap="Experimental procedure.", fig.align="center", include=TRUE, echo=FALSE}

knitr::include_graphics('Figures/expprocedure.pdf')

@@

During presentation, words in each DRM list were presented visually,\footnote{Newbury and Monaghan [@newbury2019a] found no evidence in their meta-analysis that the modality of presentation modulated the effect of sleep. However, in a set of three experiments, Fenn et al. [@fenn2009a] used auditory presentation in one of them and visual in the other two. As far as we are aware, this is the only study in the ‘Sleep x DRM’ literature that had used both modalities in the same set of experiments. Sleep appeared to have a larger effect on false memory when visual (vs. auditory) presentation was used. Given this, we opted for visual presentation in the current experiment. } in a fixed order and arranged in descending associative strength to the unpresented critical lure (see Table 2 for order). Each list began with a fixation for 1 s, followed by the first word in a list. Each word was shown for 1 s, in a lowercase black font (Arial, size 26) on a white background, and separated by a 500-ms interstimulus interval. After presentation of the final word in a list was 5 s of blank screen. List order was randomised, and each list was seen once.

@@

The free recall data were pre-processed. The first step was to remove any duplicate responses. The second was to correct all obvious spelling and typing errors to the nearest English words, defined as Levenstein distance $\leq$ 2 (e.g., *cigerette → cigarette).\footnote{If a misspelt response has more than one nearest word, the response was considered as an intrusion.} Responses with added or dropped inflectional suffixes (i.e., -s, -ed, -ing, adjectival -er) were corrected. Responses with derivational changes were considered as intrusions. For instance, one of the studied words is pollution; if a participant recalled pollutions, the plural suffix was dropped; however, if a participant recalled pollutant, this was considered as an intrusion. 
# Results of Pre-registered Analyses

@@

## Control analysis.
Payne et al. [-@payne2009a] found that participants falsely recalled more critical lures post-sleep (vs. post-wake). However, it is possible that participants simply had a greater tendency to put down more unseen words after sleep, not because sleep increases DRM false recall per se. Therefore, before addressing our key research questions, we checked if participants across groups were comparable in terms of their bias in producing unseen items. In Payne et al. [-@payne2009a], this bias was indexed via the number of intrusions (i.e., neither the studied nor the lure items), which was roughly equivalent between their Sleep and Wake groups [MSleep = 5.6 vs. MWake = 6.2; p = .60] as well as between their AM and PM-control groups [MAM = 4.1 vs. MPM = 4.1; p = .99]. To check if this is the case in our data, we used a 2 (Interval: Immediate vs. Delay) $\times$ 2 (Test Time: AM vs. PM) Poisson regression. We chose Poisson regression, as opposed to ANOVA, because the intrusion data were count data, meaning that data distribution were right-skewed and hence unsuitable for ANOVA. Figure \@ref(fig:intrusionfig) summarises the number of intrusions in each group.

```{r intrusionfig, fig.cap="Number of intrusions produced, summarised across the four groups. Each dot represents an individual participant, and the error bars represent 95\\% confidence intervals.", fig.align="center", include=TRUE, echo=FALSE}

knitr::include_graphics('Figures/intrusiongraph.pdf')

@@

The number of critical lures falsely recalled is summarised across groups in Figure \@ref(fig:lurefig).

```{r lurefig, fig.cap="Number of critical lures falsely recalled, summarised across the four groups. Each dot represents an individual participant, and the error bars represent 95\\% confidence intervals.", fig.align="center", include=TRUE, echo=FALSE}

knitr::include_graphics('Figures/luregraph.pdf')
